---
title: "Extracting Masters Data"
author: "Richard G. Gardiner"
date: "1/3/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading data and Packages

```{r}
library(tidyverse)
library(readxl)

legislative_control <- read_excel("data/Full Legislative Control.xlsx")

political_culture <- read_csv("data/political culture.csv")

professional <- read_csv("data/ncsl professionalization.csv")

ideology <- read_csv("data/pivoted_branches.csv")

legislative_partisanship <- read_csv("data/full_legislative_control.csv")

state_abb <- read_csv("state_abb.csv")
```

```{r}
years <- 1990:2018

legislative_control_long <- legislative_control %>%
  gather(`1990`, `1991`, `1992`, `1993`, `1994`, `1995`, `1996`, `1997`, `1998`, `1999`, `2000`,
         `2001`, `2002`, `2003`, `2004`, `2005`, `2006`, `2007`, `2008`, `2009`, `2010`, `2011`,
         `2012`, `2013`, `2014`, `2015`, `2016`, `2017`, `2018`,
         key = "year", value = "unified")

legislative_control_long <- legislative_control_long %>%
  filter(year >= 1995, year <= 2012)

legislative_control_long$year <- as.numeric(legislative_control_long$year)
```

# Manipulating Ideology

Need to create ideology scores for the odd years then combine them all!
```{r}
ideology <- ideology %>%
  filter(Year >= 1994)

ideology

ideology_1995 <- ideology %>%
  filter(Year == 1994) %>%
  mutate(Year = 1995)

ideology_1997 <- ideology %>%
  filter(Year == 1996) %>%
  mutate(Year = 1997)

ideology_1999 <- ideology %>%
  filter(Year == 1998) %>%
  mutate(Year = 1999)

ideology_2001 <- ideology %>%
  filter(Year == 2000) %>%
  mutate(Year = 2001)

ideology_2003 <- ideology %>%
  filter(Year == 2002) %>%
  mutate(Year = 2003)

ideology_2005 <- ideology %>%
  filter(Year == 2004) %>%
  mutate(Year = 2005)

ideology_2007 <- ideology %>%
  filter(Year == 2006) %>%
  mutate(Year = 2007)

ideology_2009 <- ideology %>%
  filter(Year == 2008) %>%
  mutate(Year = 2009)

ideology_2011 <- ideology %>%
  filter(Year == 2010) %>%
  mutate(Year = 2011)

ideology_2011

full_ideology <- rbind(ideology, ideology_1995, ideology_1997, ideology_1999, ideology_2001, ideology_2003, ideology_2005, ideology_2007,
                       ideology_2009, ideology_2011) %>%
  filter(!(State %in% "fd")) %>%
  group_by(State, Year) %>%
  arrange(State, Year)

full_ideology <- full_ideology %>%
  filter(Year > 1994)
```


## Working on Superseded datasets

```{r}
library(readxl)

short_1995 <- read_excel("data/1995 short.xlsx") # 1995 is already in long format
short_1996 <- read_excel("data/1996 short.xlsx")
short_1997 <- read_excel("data/1997 short.xlsx")
short_1998 <- read_excel("data/1998 short.xlsx")

colnames(short_1995)
colnames(short_1996)
colnames(short_1997)
colnames(short_1998)

# Getting rid of state code
short_1995 <- short_1995 %>%
  select(-state_code, -notes, -decs_yr, -decs_mo) %>%
  mutate(case_year = 1995) 

short_1996 <- short_1996 %>%
  select(-docket_n, -decs_day, -agency_r, -dec1_yr, -decs_mo) %>%
  mutate(case_year = 1996) 

short_1997 <- short_1997 %>%
  select(-decs_day, -agency_r, -dec1_yr, -decs_yr, -decs_mo) %>%
  mutate(case_year = 1997)

short_1998 <- short_1998 %>%
  select(-decs_day, -agency_r, -dec1_yr, -decs_yr, -decs_mo) %>%
  mutate(case_year = 1998)
```

1995 is already in long format.  Agency will not be included with 1995 data, amicus won't be with 1997 data.




The code below creates a data frame that has each state and their election system.
```{r}
election <- short_1995 %>%
  select(state_2, partisan, non_partisan, appointed, appoint_reelect)

election_state <- unique(election)

index <- 1:51

election_state <- cbind(election_state, index) 

election_state <- election_state %>%
  filter(!is.na(state_2)) %>%
  rename(state_abb = state_2,
         state_2 = index)

election_state
```

Assigning election type to the 3 other years

```{r}
short_1996 <- short_1996 %>%
  left_join(election_state, by = "state_2")

short_1997 <- short_1997 %>%
  left_join(election_state, by = "state_2")

short_1998 <- short_1998 %>%
  left_join(election_state, by = "state_2")
```


Now giving NA values for 1997 in classact and amicus to make the rbind work and changing name of for year of override for all groups:

```{r}
short_1997$classact <- NA
short_1997$amicus <- NA

# renaming the columns to combine all rows
short_1996 <- short_1996 %>%
  rename(override_year = YearOfOverride)

short_1997 <- short_1997 %>%
  rename(override_year = `Date of Override`)

short_1998 <- short_1998 %>%
  rename(override_year = Year_of_override)

all_years_minus_1995 <- rbind(short_1996, short_1997, short_1998)
```



Now I will try to start merging the other covariates with the `all_years_minus_1995`

```{r}
all_years_minus_1995 <- all_years_minus_1995 %>%
  left_join(political_culture, by = "state_2")

all_years_minus_1995 <- all_years_minus_1995 %>%
  left_join(professional, by = "State")

```

### Merging Covariates

```{r}
library(stringr)
all_years_minus_1995$state_abb <- toupper(all_years_minus_1995$state_abb)
state_abb$state_abb <- toupper(state_abb$state_abb)

# full_ideology <- full_ideology %>%
#   left_join(state_abb, by = c("State" = "state_abb"))
# 
# 
# ideology_legislative_control <- full_ideology %>%
#   left_join(legislative_control_long, by = c("State.y" = "State", "Year" = "year"))

all_years_minus_1995 <- all_years_minus_1995 %>%
  full_join(full_ideology, by = c("state_abb" = "State"))

# Can't do this one until I merge ideology 
all_years_minus_1995 <- all_years_minus_1995 %>%
  left_join(legislative_control_long, by = c("State", "Year" = "year"))

head(all_years_minus_1995)
```

## Filtering out rows that are unneeded

Now I need to take `all_years_minus_1995` and filter out the rows for cases in which there is an override, and the year is the past the override.  This will likely have some problems with the NAs, so will need to experiment.  I think temporarily, I will change the NA's to 2020 and then after filtering, switch them back to NA.

```{r}

all_years_minus_1995$override_year <- replace_na(all_years_minus_1995$override_year, 2020)


all_years_minus_1995 <- all_years_minus_1995 %>%
  filter(Year <= override_year)

all_years_minus_1995$override_year <- ifelse(all_years_minus_1995$override_year == 2020, NA, all_years_minus_1995$override_year)
```

Looks like it worked, but now I have found another problem, rows with `Year` that are before the `case_year`.  Need to filter out those:
```{r}
all_years_minus_1995 <- all_years_minus_1995 %>%
  filter(Year >= case_year)
```

## Fixing the override variable

The override variable needs to be 0 for only the exact year in which it was overriden.  So `all_years_minus_1995$Override` should be around 300ish, not over 1000.  The following code fixes it.
```{r}
table(all_years_minus_1995$Override)
all_years_minus_1995$Override <- ifelse(all_years_minus_1995$Override == 1 & all_years_minus_1995$override_year > all_years_minus_1995$Year,
                                  0, 
                                  all_years_minus_1995$Override)
table(all_years_minus_1995$Override)
```

# 1995


This is taking the data from the masters and getting it ready.
```{r}

library(haven)
march_13_full_dataset <- read_dta("march 13 full dataset.dta")
head(march_13_full_dataset)

march_13_full_dataset %>%
  filter(new_superseded == 1)

march_13_full_dataset %>%
  filter(docket == 1940087)

short_1995_v2 <- march_13_full_dataset %>%
  select(masterid, agency, state_2, new_superseded, docket,
         amicus, decs_yr) %>%
  mutate(classact = NA,
         case_year = 1995) %>%
  rename(state_abb = state_2,
         cutt_off_year = decs_yr, # this is differnet than override_year because this is the dataset already to go
         Override = new_superseded) %>%  
  left_join(election_state, by = "state_abb")
```







## Adding covariates for 1995
```{r}

short_1995_v2 <- short_1995_v2 %>%
  left_join(political_culture, by = "state_2")

short_1995_v2 <- short_1995_v2 %>%
  left_join(professional, by = "State")

short_1995_v2$state_abb <- toupper(short_1995_v2$state_abb)


short_1995_v2 <- short_1995_v2 %>%
  rename(Year = cutt_off_year) %>%
  full_join(full_ideology, by = c("state_abb" = "State", "Year" = "Year"))

short_1995_v2 <- short_1995_v2 %>%
  left_join(legislative_control_long, by = c("State", "Year" = "year"))
```

# Merging the 4 years

Need to take out docket from short_1995_v2 and take out override_year from all_years_minus_1995.  Then I should be able to do a simple rbind

```{r}
final_1995 <- short_1995_v2 %>%
  select(-docket, -X5)

final_all_other_years <- all_years_minus_1995 %>%
  select(-override_year, -X5)

all_years <- rbind(final_1995, final_all_other_years)

all_years
```

## Merging the Full Leigslative Control variable
```{r}
all_years <- all_years %>%
  left_join(legislative_partisanship, by = c("Year", "State"))
```

# Write the file

Now I simply write a csv and start a new rmd file to begin EDA:

```{r}
# write_csv(all_years, "data/all_years_all_variables.csv", na = "")
```

