---
title: "EDA"
author: "Richard G. Gardiner"
date: "1/16/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data and Packages

```{r}
library(tidyverse)
library(usmap)
library(maps)
```

```{r}
full_dataset <- read_csv("data/all_years_all_variables.csv")
head(full_dataset)
```

First, let's just look at the basic characteristics of the data by looking at number of cases per state and year
```{r}
full_dataset %>%
  group_by(masterid, state_abb, case_year) %>%
  summarise(cases = n()) %>%
  ungroup() %>%
  ggplot(aes(case_year)) +
  geom_bar()
```


```{r}
all_states <- map_data("state")


full_dataset$State <- tolower(full_dataset$State)


full_dataset2 <- full_dataset %>%
  distinct(masterid, .keep_all = TRUE) %>%
  group_by(State) %>%
  count()

all_states2 <- all_states %>%
  left_join(full_dataset2, by = c("region" = "State"))

ggplot() +
  geom_polygon(data = all_states2, aes(x = long, y = lat, group = group, fill = all_states2$n)) +
  scale_fill_gradient(low = "white", high = "red") +
  labs(fill = "Number of Cases") +
  theme_void()
```


## Overrides Only

The only problem with the chart below is that this doesn't take into account that way more states have elections than appointments.
```{r}

overrides <- full_dataset %>%
  filter(Override == 1)

overrides$election <- ifelse(overrides$partisan == 1, "partisan",
                             ifelse(overrides$non_partisan == 1, "nonpartisan",
                             ifelse(overrides$appointed == 1, "appointed",
                             "merit")))

overrides %>%
  group_by(election) %>%
  summarise(total = sum(Override)) %>%
  ggplot() +
  geom_col(aes(x = election, y = total)) +
  ylab("Total Number of Overrides") +
  xlab("Selection System")
```

This chart does show, however, the distribution of values versus the graph above that looks at raw numbers.
```{r}
overrides %>%
  group_by(election, case_year) %>%
  summarise(total = sum(Override)) %>%
  ggplot() +
  geom_boxplot(aes(x = election, y = total)) +
  ylab("Total Number of Overrides") +
  yxlab("Selection System")
```

Distance between case decided and override. Initial support found.
```{r}
overrides$time_to_override <- overrides$Year - overrides$case_year
overrides$court_lower <- abs(overrides$judicial - overrides$lower)
overrides$court_upper <- abs(overrides$judicial - overrides$upper)



ggplot() +
  geom_bar(data = overrides, aes(x = time_to_override, fill = election)) +
  xlab("Years between Decision and Override") +
  ylab("Count")
```

Ideological distance between court and branches when overridden
```{r}
a <- ggplot() +
  geom_histogram(data = overrides, aes(x = court_lower)) +
  xlab("Distance between Court and Lower Chamber") +
  ylab("Count")

b <- ggplot() +
  geom_histogram(data = overrides, aes(x = court_upper)) +
  xlab("Distance between Court and Lower Chamber") +
  ylab("Count")


gridExtra::grid.arrange(a, b)
```

I think a graph like this is important, but not going to happen right now.  If I do want a graph like this, then I will need to do something like put missing values at the median for the state
```{r}
overrides %>%
  group_by(election, case_year) %>%
  ggplot() +
  geom_boxplot(aes(x = election, y = court_lower))
```



```{r}
overrides_by_state <- overrides %>%
  group_by(State) %>%
  count()

override_state_map <- all_states %>%
  left_join(overrides_by_state, by = c("region" = "State"))

ggplot() +
  geom_polygon(data = override_state_map, aes(x = long, y = lat, group = group, fill = n)) +
  scale_fill_gradient(low = "white", high = "red") +
  labs(fill = "Number of Overrides") +
  theme_void()
```

