---
title: "EDA"
author: "Richard G. Gardiner"
date: "1/16/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data and Packages

```{r}
library(tidyverse)
library(usmap)
library(maps)

theme_set(theme_light())

full_dataset <- read_csv("data/all_years_all_variables.csv")	
head(full_dataset)
```

```{r}
full_dataset %>%
  distinct(masterid, state_abb, case_year) %>%
  count(case_year) %>%
  ggplot(aes(x = case_year, y = n)) +
  geom_col() +
  labs(x = "Year of Decision", y = "Number of Cases per Year")


# ggsave("figures/cases per year.jpeg")
```


```{r}
all_states <- map_data("state")


full_dataset$State <- tolower(full_dataset$State)


full_dataset2 <- full_dataset %>%
  distinct(masterid, .keep_all = TRUE) %>%
  group_by(State) %>%
  count()

all_states2 <- all_states %>%
  left_join(full_dataset2, by = c("region" = "State"))

ggplot() +
  geom_polygon(data = all_states2, aes(x = long, y = lat, group = group, fill = all_states2$n)) +
  scale_fill_gradient(low = "white", high = "red") +
  labs(fill = "Number of Cases") +
  theme_void()

# ggsave("figures/cases by state.jpeg")
```


## Overrides Only

The only problem with the chart below is that this doesn't take into account that way more states have elections than appointments.
```{r}

overrides <- full_dataset %>%
  filter(Override == 1)

overrides$election <- ifelse(overrides$partisan == 1, "partisan",
                             ifelse(overrides$non_partisan == 1, "nonpartisan",
                             ifelse(overrides$appointed == 1, "appointed",
                             "merit")))
overrides$culture <- ifelse(overrides$Traditionalistic == 1, "traditionalistic",
                            ifelse(overrides$Moralistic == 1, "moralistic",
                                   "individualistic"))

ggplot(data = overrides) +
  geom_bar(aes(case_year)) +
  xlab("Year when Case was Decided") +
  ylab("Number of Overrides")

# ggsave("figures/overrides by year case decided.jpeg")

table(overrides$unified)
table(full_dataset$unified)
```

```{r}
ggplot(data = overrides) +
  geom_bar(aes(x = Year, fill = as.factor(case_year))) +
  xlab("Year when Case was Overridden") +
  ylab("Number of Overrides") +
  labs(fill = "Year Case Decided")
  

# ggsave("figures/Year overridden by Case Year.jpeg")
```


```{r}
overrides %>%
  group_by(election) %>%
  summarise(total = sum(Override)) %>%
  ggplot() +
  geom_col(aes(x = election, y = total)) +
  ylab("Total Number of Overrides") +
  xlab("Selection System")

# ggsave("figures/overrides by selection system.jpeg")
```

This chart does show, however, the distribution of values versus the graph above that looks at raw numbers.
```{r}
# overrides %>%
#   group_by(election, case_year) %>%
#   summarise(total = sum(Override)) %>%
#   ggplot() +
#   geom_boxplot(aes(x = election, y = total)) +
#   ylab("Total Number of Overrides") +
#   xlab("Selection System")


overrides %>%
  group_by(state_abb, election, case_year) %>%
  summarise(total = sum(Override)) %>%
  ungroup() %>%
  group_by(state_abb, election) %>%
  summarise(mean = mean(total)) %>%
  ggplot() +
  geom_boxplot(aes(x = election, y = mean)) +
  ylab("Total Number of Overrides") +
  xlab("Selection System") 

# ggsave("figures/distribution of overrides by selection system.jpeg")
```

Distance between case decided and override. Initial support found.
```{r}
overrides$time_to_override <- overrides$Year - overrides$case_year
overrides$court_lower <- abs(overrides$judicial - overrides$lower)
overrides$court_upper <- abs(overrides$judicial - overrides$upper)

ggplot() +
  geom_bar(data = overrides, aes(x = time_to_override)) +
  xlab("Years between Decision and Override") +
  ylab("Number of Overrides")

# ggsave("figures/year between case and override.jpeg")

ggplot() +
  geom_bar(data = overrides, aes(x = time_to_override, fill = election)) +
  xlab("Years between Decision and Override") +
  ylab("Number of Overrides") +
  labs(fill = "Selection System")

# ggsave("figures/year between case state and override.jpeg")
```

```{r}
ggplot() +
  geom_density(data = overrides, aes(x = time_to_override, color = election)) +
  xlab("Years between Decision and Override") +
  ylab("Density")

# ggsave("figures/Density Plot of Years by Selection Type.jpeg")
```

I think a graph like this is important, but not going to happen right now.  If I do want a graph like this, then I will need to do something like put missing values at the median for the state
```{r}
overrides %>%
  group_by(election, case_year) %>%
  ggplot() +
  geom_boxplot(aes(x = election, y = court_lower))
```



```{r}
overrides_by_state <- overrides %>%
  group_by(State) %>%
  count()

override_state_map <- all_states %>%
  left_join(overrides_by_state, by = c("region" = "State"))

ggplot() +
  geom_polygon(data = override_state_map, aes(x = long, y = lat, group = group, fill = n)) +
  scale_fill_gradient(low = "white", high = "red") +
  labs(fill = "Number of Overrides") +
  theme_void()

# ggsave("figures/map of overrides.jpeg")
```


### Smaller variables

These would include agency, amicus, moralistic, traditionalistic


```{r}
smaller_variable_eda <- full_dataset %>%
  distinct(masterid, .keep_all = TRUE)
```


```{r}
table(overrides$agency) # 88 means not relevant
table(smaller_variable_eda$agency) # full dataset
3263/22835 # about 14%
52/338 # about 15%
```

```{r, amicus}
table(overrides$amicus) # overrides
table(smaller_variable_eda$amicus)
46/279 # 15%
1642/19712 # 8%
```


```{r}
ggplot(data = overrides) +
  geom_bar(aes(culture)) +
  ylab("Number of Overrides") +
  labs(x = "")

# ggsave("figures/political culture.jpeg")
```

```{r}
ggplot(data = overrides) +
  geom_density(aes(x= time_to_override, color = as.factor(culture))) +
  labs(color = "Political Culture") +
  xlab("Time to Override")

# ggsave("figures/Density plot by Culture.jpeg")

overrides %>%
  group_by(culture) %>%
  summarise(mean = mean(time_to_override),
            standard_dev = sd(time_to_override))
```

Moralistic was 120 overrides, Traditionalistic had 150 and individualualistic had 154.

The total number of cases for each were: 9121 for Moralistic, 8731 for traditionalistic, and 10359 for individualistic.  Seems like there is a clear trend where traditionalistic has higher rate of overrides.

Unified:
```{r}
table(smaller_variable_eda$unified)

table(overrides$unified)

table(full_dataset$unified)

105/6094

311/21318
```

```{r}
table(overrides$unified)
```

Professionalization:

The first plot really isn't telling us much, givne that the bulk of observations are simply with level # 3.  More of a demonimator problem.
```{r}
ggplot(data = overrides) +
  geom_bar(aes(Professionalization)) +
  labs(y = "Number of Overrides",
       x = "Level of Professionalization")

table(full_dataset$Professionalization)

# ggsave("figures/professionalization.jpeg")

ggplot() +
  geom_density(data = overrides, aes(x = time_to_override, 
                                     color = as.factor(Professionalization))) +
  xlab("Years between Decision and Override") +
  ylab("Density") +
  labs(color = "Level of Professionalization")

# ggsave("figures/Professionalization and time to override.jpeg")
```

Salience:
```{r}
table(smaller_variable_eda$case_year, smaller_variable_eda$amicus)

table(smaller_variable_eda$amicus)

table(smaller_variable_eda$amicus, smaller_variable_eda$Override)

41/19652

7/1632
```

## New Ideology with Amelia

I am uploading the data from the amelia dataset that will hopefully help with the data problem.

```{r}
amelia <- read_csv("data/Amelia_data1.csv")
```

Ideological distance between court and branches when overridden
```{r}

a_overrides <- amelia %>%
  filter(Override == 1) %>%
  mutate(time_to_override = Year - case_year)



a_overrides %>%
  group_by(state_abb, court_lower_dist) %>%
  count() %>%
  ggplot() +
  geom_point(aes(x = court_lower_dist, y = n)) +
  xlab("Distance between Court and Lower Chamber") +
  ylab("Count of Overrides") +
  ylim(0,10)

# ggsave("figures/lower ideology and overrides.jpeg")


a_overrides %>%
  group_by(state_abb, court_upper_dist) %>%
  count() %>%
  ggplot() +
  geom_point(aes(x = court_upper_dist, y = n)) +
  xlab("Distance between Court and Upper Chamber") +
  ylab("Count of Overrides") +
  ylim(0,10)


# ggsave("figures/upper ideology and overrides.jpeg")
```


```{r}
ggplot(data = a_overrides, aes(x = court_upper_dist, y = time_to_override, 
                               color = as.factor(appointed))) +
  geom_point() +
  stat_smooth(method = "lm") +
  xlab("Distance between Court and Upper Chamber") +
  ylab("Time between Decision and Override") +
  scale_color_manual(name = "Election Status",
                     breaks = c(0, 1),
                     values = c("red", "blue"),
                     labels = c("Elected", "Appointed"))

# ggsave("figures/Upper Ideology and Time.jpeg")

ggplot(data = a_overrides, aes(x = court_lower_dist, y = time_to_override,
                               color = as.factor(appointed))) +
  stat_smooth(method = "lm") +
  geom_point() +
  xlab("Distance between Court and Lower Chamber") +
  ylab("Time between Decision and Override") +
  scale_color_manual(name = "Election Status",
                     breaks = c(0, 1),
                     values = c("red", "blue"),
                     labels = c("Elected", "Appointed"))

# ggsave("figures/Lower Ideology and Time.jpeg")
```



## New Ideology Dichotomous


```{r}

full_dataset %>%
  group_by(Year) %>%
  summarize(sum = sum(ideologically_similar))

full_dataset %>%
  filter(!is.na(ideologically_similar)) %>%
  mutate(Year = as.Date(as.character(Year), "%Y"),
         appointed = ifelse(appointed == 1, "Appointed", "Elected")) %>%
  group_by(appointed, Year) %>%
  summarize(mean_ideology_similar = mean(ideologically_similar)) %>%
  ggplot(aes(x = Year, y = mean_ideology_similar, color = as.factor(appointed))) +
  geom_line() +
  labs(y = "Proportion of Cases with Ideologically Disimilar Branches") +
  scale_color_discrete(name = "Selection System") +
  theme_light() +
  ylim(c(0,1))

# ggsave("figures/Ideology over time.jpeg")
```

```{r}
overrides %>%
  filter(!is.na(ideologically_similar)) %>%
  ggplot() +
  geom_density(aes(x = time_to_override, color = as.factor(ideologically_similar))) +
  xlab("Years between Decision and Override") +
  ylab("Density") +
  labs(color = "Ideologically Similar") +
  scale_color_discrete(labels = c("No", "Yes"))
  theme_light() 

# ggsave("figures/ideology and tiem to override.jpeg")
```

```{r}
table(full_dataset$ideologically_similar)

overrides %>%
  filter(!is.na(ideologically_similar)) %>%
  ggplot(aes(x = as.factor(ideologically_similar), fill = election)) +
  geom_bar() +
  labs(x = "Ideologically Similar",
       fill = "Selection System",
       y = "Number of Overrides") +
  scale_x_discrete(labels = c("No", "Yes")) +
  theme_light()

# ggsave("figures/Overrides by Ideological similarity.jpeg")
```


