---
title: "Model"
author: "Richard G. Gardiner"
date: "2/20/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data and Packages

When you start looking at the model, look at [David's Tidy Tuesday project](https://mobile.twitter.com/drob/status/1082318029168566273?s=21). Also check out [this post](https://hockey-graphs.com/2019/01/07/quantifying-differences-between-the-regular-season-and-playoffs-using-survival-analysis/?blogsub=confirming#subscribe-blog).  Lastly, I like [this link](https://rpubs.com/daspringate/survival)

```{r, packages}
library(tidyverse)
library(survival)
library(survminer)
library(broom)
library(scales)
library(devtools)
library(ggfortify)
library(survivalAnalysis)
library(Amelia)
library(frailtypack)

theme_set(theme_light())
```

```{r}
dataset <- read_csv("data/all_years_all_variables.csv")
head(dataset)

dataset <- dataset %>%
  mutate(Year = Year + 0.01,
         Override = ifelse(Override == 109, 0, Override))



dataset %>%
  group_by(masterid) %>%
  count(masterid)

dataset <- dataset %>%
  mutate(appointed_similar = appointed * ideologically_similar)

control_dataset <- dataset %>%
  filter(agency != 88) %>%
  mutate(Individualistic = ifelse(Traditionalistic == 0 & Moralistic == 0, 1, 0))
```

## Amelia 

This uses the data from Amelia to fill in missing data.  
```{r Amelia}
# filled_data <- amelia(dataset, idvars = c("state_abb", "State"))
# 
# write.amelia(filled_data, file.stem = "Amelia_data", format = "csv")
# 
# dataset <- read_csv("data/Amelia_data1.csv")
# 
# dataset
```


Here I am adding the absolute distance variables:

```{r}
# dataset$court_gov_dist <- abs(dataset$judicial - dataset$governor)
# dataset$court_lower_dist <- abs(dataset$judicial - dataset$lower)
# dataset$court_upper_dist <- abs(dataset$judicial - dataset$upper)
```



## Initial Analysis (not Amelia)

```{r}
# Model
election_cluster <- coxph(formula = Surv(case_year, Year, Override) ~ appointed + cluster(state_2),
                 data = dataset)
summary(election_cluster)

# graph prep
elections_cluster_fit <- survfit(formula = Surv(case_year, Year, Override) ~ appointed + cluster(state_2),
                 data = dataset)
# graph
tidy(elections_cluster_fit) %>%
  ggplot(aes(time, estimate, color = strata)) +
  geom_line() +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .2) +
  scale_color_discrete(name = "Selection System",
                       breaks = c("appointed=1", "appointed=0"),
                       labels = c("Appointed", "Elected")) +
  theme_light() +
  labs(y = "Estimated Likelihood of Survival",
       x = "Year")

# ggsave("figures/Selection Survival.jpeg")

cox.zph(election_cluster)
```

```{r}
ideology_cluster <- coxph(formula = Surv(case_year, Year, Override) ~ appointed +
                ideologically_similar + cluster(state_2),
                data = dataset)

summary(ideology_cluster)

ideology_cluster_fit <- survfit(formula = Surv(case_year, Year, Override) ~ appointed +
                ideologically_similar + cluster(state_2),
                data = dataset)


tidy(ideology_cluster_fit) %>%
  mutate(strata = if_else(strata == "appointed=0, ideologically_similar=1", "Elected and Similar",
                          if_else(strata == "appointed=0, ideologically_similar=0", "Elected and Dissimilar",
                                  if_else(strata == "appointed=1, ideologically_similar=0", 
                                          "Appointed and Dissimilar", "Appointed and Similar")))) %>%
  ggplot(aes(time, estimate, color = strata)) +
  geom_line() +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .2) +
  facet_wrap(~strata) +
  theme_light() +
  labs(y = "Estimated Likelihood of Survival",
       x = "Year",
       color = "Selection System and \nIdeological Similarity")

# ggsave("figures/Selection and Ideology survival.jpeg")

cox.zph(ideology_cluster)
```

```{r}
# creating interaction term



interactions_cluster <- coxph(formula = Surv(case_year, Year, Override) ~ 
                                appointed + ideologically_similar + appointed_similar +
                                + cluster(state_2),
                data = dataset)
summary(interactions_cluster)

interactions_cluster_fit <- survfit(formula = Surv(case_year, Year, Override) ~ appointed +
                              ideologically_similar + appointed_similar + cluster(state_2),
                data = dataset)

tidy(interactions_cluster_fit) %>%
  mutate(strata = if_else(strata == "appointed=0, ideologically_similar=0, appointed_similar=0", 
                          "Elected and Dissimilar",
                          if_else(strata == "appointed=0, ideologically_similar=1, appointed_similar=0",
                                  "Elected and Similar",
                                  if_else(strata == 
                                            "appointed=1, ideologically_similar=1, appointed_similar=1", 
                                          "Appointed and Similar", "Appointed and Dissimilar")))) %>%
  filter(strata %in% c("Elected and Dissimilar", "Elected and Similar", "Appointed and Similar",
                       "Appointed and Dissimilar")) %>%
  ggplot(aes(time, estimate, color = strata)) +
  geom_line() +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .2) +
  facet_wrap(~strata) +
  theme_light() +
  labs(y = "Estimated Likelihood of Survival",
       x = "Year",
       color = "Interaction between Selection System \nand Ideological Similarity")  

# ggsave("figures/Interaction Survival.jpeg")

cox.zph(interactions_cluster)
```

```{r}
controls_cluster <- coxph(formula = Surv(case_year, Year, Override) ~ appointed +
                ideologically_similar + appointed_similar + classact + Individualistic + Traditionalistic +
                Professionalization + unified + cluster(state_2),
                data = control_dataset)

summary(controls_cluster)

controls_cluster_fit <- survfit(formula = Surv(case_year, Year, Override) ~ appointed +
                ideologically_similar + appointed_similar + classact + Individualistic + Traditionalistic +
                Professionalization + unified + cluster(state_2),
                data = control_dataset)

tidy(controls_cluster_fit) %>%
  mutate(strata = if_else(strata == "appointed=0, ideologically_similar=0, appointed_similar=0, classact=FALSE, Individualistic=0, Traditionalistic=0, Professionalization=1, unified=1", "elected, similar, not professional",
                           if_else(strata == "appointed=0, ideologically_similar=0, appointed_similar=0, classact=FALSE, Individualistic=0, Traditionalistic=0, Professionalization=5, unified=1", "elected, similar, most professional", strata))) %>%
  filter(strata %in% c("elected, similar, not professional", "elected, similar, most professional",
                       "appointed, similar, not professional", "appointed, similar, most professional")) %>%
  ggplot(aes(time, estimate, color = strata)) +
  geom_line() +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .2) +
  facet_wrap(~strata) +
  theme_light() +
  labs(y = "Estimated Likelihood of Survival",
       x = "Year",
       color = "Selection System and Professionalization")  

# ggsave("figures/survival controls.jpeg")

cox.zph(controls_cluster)

print(tidy(controls_cluster_fit$strata))
```








## Using videos in attempt to better understand survival models

Creating survival object
```{r}
mySurv <- Surv(time = dataset$case_year, time2 = dataset$Year, event = dataset$Override)
head(mySurv) # plus sign means that it is centered
```


Baseline with no predictor
```{r}
fit <- survfit(mySurv ~ 1)

fit

plot(fit, ylim = c(0.9, 1), xlim = c(1995, 2013)) # kaplan meier curve
?plot
```

These Kaplan Meier tests/curves are nice, but have limitations.  It can only incorporate categorical variables.  It becomes difficult when you have many x variables (controlling becomes too hard), and there is no good way to summarize relationships (no betas) given that it is non-parametric
```{r}
states <- survfit(mySurv ~ dataset$State)

states

# kaplan meier curve
plot(fit, col = "red", lwd = 8, ylim = c(0.992, 1), xlim = c(1995, 2013), conf.int = FALSE) # baseline
lines(states) # state model


# trying to run a log rank test (like a chi-square test), but it is giving me an error
# survdiff(mySurv ~ dataset$state_abb) 
```

Elections
```{r}
elections <- survfit(mySurv ~ dataset$appointed)

plot(elections, ylim = c(0.998, 1), xlim = c(1995, 2013), col = c("blue", "red"))
```





### Cox proportional hazard ratio analysis

- Survival analysis allowing inclusing of multiple predictors
- Proportional Hazard assumption
    - assumes a constant hazard ratio over time
    - Groups compared diverge in the KM plot at the same rate (parallel)
- Coefficients represent the increase in the log Hazard Rate for one unit increase in predictor with all other covariates constant
- e^beta represents the Hazard rate for one unit increase in predictor with all others constant
- beta < 0 means that increasing x is associated with lower risk (hazard) and longer survival times
- to test hypothesis of beta = 0, we can use a Wald or LR test

```{r}
cox_states <- coxph(mySurv ~ dataset$State)

summary(cox_states)

tidy(cox_states) %>%
  mutate(term = str_remove_all(term, "dataset\\$State"),
         estimate2 = exp(estimate),
         significance = ifelse(estimate > 0 & p.value < 0.05, "positive",
                               ifelse(estimate > 0 & p.value < 0.05, "negative",
                                      "not significant")),
         term = fct_reorder(term, estimate)) %>%
  # filter(p.value < 0.05) %>%
  filter(term != "New Mexico") %>%
  ggplot(aes(x = term, y = estimate, color = significance)) +
  geom_point() +
  geom_pointrange(aes(ymin = conf.low, ymax = conf.high)) +
  scale_color_manual(values = c("gray60", "green2")) +
  coord_flip() +
  theme_minimal()

tidy(cox_states) %>%
  mutate(term = str_remove_all(term, "dataset\\$State"),
         estimate2 = exp(estimate),
         significance = ifelse(estimate > 0 & p.value < 0.05, "positive",
                               ifelse(estimate > 0 & p.value < 0.05, "negative",
                                      "not significant")),
         term = fct_reorder(term, conf.low)) %>%
  filter(p.value < 0.05) %>%
  filter(term != "New Mexico") %>%
  ggplot(aes(x = term, y = estimate)) +
  geom_point() +
  geom_pointrange(aes(ymin = conf.low, ymax = conf.high)) +
  scale_color_manual(values = c("green2")) +
  coord_flip() +
  theme_minimal()
```


Hazard is the probability that you die now, given that you are alive (probability of being overriden, given that you haven't been overridden right now).

A hazard ratio is just like an odds-ratio.  The cox proportional can't predict the hazard for a given person with a given set of values at a given moment.  We can only COMPARE people, because we don't know the value of the baseline hazard


WHAT IF I SHOULD BE SUBTRACTING THE YEAR? ehh.. Really doesn't matter
```{r}
mySurv2 <- Surv(time = (dataset$case_year - 1994), time2 = (dataset$Year - 1994), event = dataset$Override)

fitTest <- survfit(mySurv2 ~ 1)

fitTest

plot(fitTest) # kaplan meier curve
```











## Frailty using the survival package (MAIN RESULTS NOW)

Notes from https://openaccess.leidenuniv.nl/bitstream/handle/1887/11456/01.pdf?sequence=6: 

Vaupel et al. (1979) introduced the term frailty and used it in univariate survival models. Clayton (1978) promoted the model by its application to multivariate situation on chronic disease incidence in families. A random effect model takes into account the effects of unobserved or unobservable heterogeneity, caused by different sources. The random effect, called frailty and denoted here by Z is the term that describes the common risk or the individual heterogeneity, acting as a factor on the hazard function. Two categories of frailty models can be pointed out. The first one is the class of univariate frailty models that consider univariate survival times. The second one is the class of multivariate frailty models that take into account multivariate survival times.

*Univariate frailty models*

Univariate frailty models take into account that the population is not homogeneous. Heterogeneity may be explained by covariates, but when important covariates have not been observed, this leads to unobserved heterogeneity. Vaupel et al. (1979) introduced univariate frailty models (with a gamma distribution) into survival analysis to account for unobserved heterogeneity or missing covariates in the study population. The idea is to suppose that different patients possess different frailties and patients more ”frail” or ”prone” tend to have the event earlier that those who are less frail. 

*1.2.3 Multivariate frailty models*

A very common situation in survival analysis is clustered or repeated data. Clustered data are for instance data where individuals are divided in groups likes family or study centres. Repeated data are seen in case of longitudinal data, concerning multiple recurrences of an event for the same individual. The difficulty of working with this kind of data is due to the dependence of individuals within groups, or repeated measures within individuals. The dependence usually arises because individuals in the same group are related to each other or because of the recurrence of an event for the same individual. Multivariate frailty models have been used frequently for modelling dependence in multivariate time-to-event data (Clayton, 1978; Hougaard, 2000; Oakes, 1982a; Yashin et al., 1995). The aim of the frailty is to take into account the presence of the correlation between the multivariate survival times.

Notes from https://www.ncbi.nlm.nih.gov/pubmed/9385105 (houggard):

A frailty model is a random effects model for time variables, where the random effect (the frailty) has a multiplicative effect on the hazard. It can be used for univariate (independent) failure times, i.e. to describe the influence of unobserved covariates in a proportional hazards model. More interesting, however, is to consider multivariate (dependent) failure times generated as conditionally independent times given the frailty. 


Other notes: by not considering the fact that there may be clusters in the data (states), we are not accounting for the heterogeneity inherent in the data. This should increase our ability to find a true relationship.

Frailty models are appropriate for analyses that have clusetered survival data (http://www.vetstat.ugent.be/workshop/Janssen2005/frailty.pdf)


```{r}
frailty1 <- coxph(Surv(time = (dataset$case_year - 1994), time2 = (dataset$Year - 1994), 
                       event = dataset$Override) ~ frailty(dataset$state_abb))


summary(frailty1)

surv_frailty1 <- survfit(Surv(time = (dataset$case_year - 1994), 
                              time2 = (dataset$Year - 1994), 
                              event = dataset$Override) ~ 
                           dataset$appointed + frailty(dataset$state_abb))



surv_frailty1 %>%
  tidy() %>%
  ggplot(aes(time, estimate, color = strata)) +
  geom_line(show.legend = FALSE) +
  labs(x = "Time", y = "Estimate")

ggsave("figures/frailty_basic.png")
```

```{r}
frailty2 <- coxph(Surv(time = (dataset$case_year - 1994), 
                              time2 = (dataset$Year - 1994), 
                              event = dataset$Override) ~ 
                           dataset$appointed +
                           frailty(dataset$state_abb))

summary(frailty2)

surv_frailty2 <- survfit(Surv(time = (dataset$case_year - 1994), 
                              time2 = (dataset$Year - 1994), 
                              event = dataset$Override) ~ 
                           dataset$appointed +
                           frailty(dataset$state_abb))

frailty2_data <- surv_frailty2 %>%
  tidy() %>% 
  mutate(appointed = ifelse(str_detect(strata, "appointed=0") == TRUE,
                            "Elected", "Appointed")) %>%
  group_by(time, appointed) %>%
  summarise(estimate = mean(estimate),
            conf.high = mean(conf.high),
            conf.low = mean(conf.low))

frailty2_data %>%
  ggplot(aes(time, estimate, color = appointed)) +
  geom_line() +
  labs(color = "Appointed", x = "Time", y = "Estimate")

surv_frailty2 %>%
  tidy() %>% 
  mutate(appointed = ifelse(str_detect(strata, "appointed=0") == TRUE,
                            "Elected", "Appointed")) %>%
  ggplot(aes(time, estimate, color = appointed)) +
  geom_line(aes(group = strata)) +
  geom_smooth(data = frailty2_data, lwd = 1, color = "gray60", se = FALSE) +
  labs(x = "Time", y = "Estimate", color = "Retention System",
       caption = "Gray line is the group average") +
  facet_wrap(~appointed) 

ggsave("figures/frailty_retention.png")
```



```{r}
frailty3 <- coxph(Surv(time = (dataset$case_year - 1994), 
                              time2 = (dataset$Year - 1994), 
                              event = dataset$Override) ~ 
                           dataset$appointed + dataset$ideologically_similar +
                           frailty(dataset$state_abb))

summary(frailty3)

surv_frailty3 <- survfit(Surv(time = (dataset$case_year - 1994), 
                              time2 = (dataset$Year - 1994), 
                              event = dataset$Override) ~ 
                           dataset$appointed + dataset$ideologically_similar +
                           frailty(dataset$state_abb))

frailty3_data <- surv_frailty3 %>%
  tidy() %>% 
  mutate(appointed = ifelse(str_detect(strata, "appointed=0") == TRUE,
                            "Elected", "Appointed")) %>%
  group_by(time, appointed) %>%
  summarise(estimate = mean(estimate),
            conf.high = mean(conf.high),
            conf.low = mean(conf.low))

frailty3_data %>%
  ggplot(aes(time, estimate, color = appointed)) +
  geom_line() +
  labs(color = "Appointed", x = "Time", y = "Estimate")

surv_frailty3 %>%
  tidy() %>%
  dplyr::select(strata) %>%
  mutate(strata = str_remove_all(strata, "frailty.*")) %>%
  count(strata)

surv_frailty3 %>%
  tidy() %>% 
  mutate(appointed = ifelse(str_detect(strata, "appointed=0") == TRUE,
                            "Elected", "Appointed"),
         ideology = ifelse(str_detect(strata, "ideologically_similar=1"),
                                      "similar", "dissimilar"),
         final = paste(appointed, ideology)) %>%
  ggplot(aes(time, estimate, color = final)) +
  geom_line(aes(group = strata)) +
  geom_smooth(data = frailty3_data, lwd = 1, color = "gray60", se = FALSE) +
  labs(x = "Time", y = "Estimate", color = "Retention System",
       caption = "Gray line is the group average") +
  facet_wrap(~final) 

ggsave("figures/frailty_ideology.png")
```


```{r}
frailty4 <- coxph(Surv(time = (dataset$case_year - 1994), 
                              time2 = (dataset$Year - 1994), 
                              event = dataset$Override) ~ 
                           dataset$appointed + dataset$ideologically_similar +
                           dataset$appointed_similar +
                           frailty(dataset$state_abb))

summary(frailty4)

surv_frailty4 <- survfit(Surv(time = (dataset$case_year - 1994), 
                              time2 = (dataset$Year - 1994), 
                              event = dataset$Override) ~ 
                           dataset$appointed + dataset$ideologically_similar +
                           dataset$appointed_similar +
                           frailty(dataset$state_abb))

surv4_data <- surv_frailty4 %>%
  tidy() %>% 
  mutate(appointed = ifelse(str_detect(strata, "appointed=0") == TRUE,
                            "Elected", "Appointed"),
         ideological_dissimilar = ifelse(str_detect(strata, "ideologically_similar=0") == TRUE,
                                         "Dissimilar", "Similar")) %>%
  group_by(time, appointed, ideological_dissimilar) %>%
  summarise(estimate = mean(estimate),
            conf.high = mean(conf.high),
            conf.low = mean(conf.low)) %>%
  mutate(final = paste(appointed, ideological_dissimilar))
  

surv4_data %>%
  ggplot(aes(time, estimate, color = appointed)) +
  geom_line() +
  labs(color = "Appointed", x = "Time", y = "Estimate") +
  facet_wrap(~ideological_dissimilar)



surv_frailty4 %>%
  tidy() %>% 
  mutate(appointed = ifelse(str_detect(strata, "appointed=0") == TRUE,
                            "Elected", "Appointed"),
         ideological_dissimilar = ifelse(str_detect(strata, "ideologically_similar=0") == TRUE,
                                         "Dissimilar", "Similar"),
         final = paste(appointed, ideological_dissimilar)) %>%
  ggplot(aes(time, estimate, color = final)) +
  geom_line(aes(group = strata)) +
  geom_smooth(data = surv4_data, lwd = 1, color = "gray60", se = FALSE) +
  labs(x = "Time", y = "Estimate", color = "Retention System",
       caption = "Gray line is the group average") +
  facet_wrap(~final) 

ggsave("figures/frailty-interaction.png")
```





```{r}
controls_cluster_frail <- coxph(formula = Surv(case_year, Year, Override) ~ appointed +
                ideologically_similar + appointed_similar + classact + Individualistic + Traditionalistic +
                Professionalization + unified + frailty(state_2),
                data = control_dataset)

summary(controls_cluster_frail)

controls_cluster_fit_frail <- survfit(formula = Surv(case_year, Year, Override) ~ appointed +
                ideologically_similar + appointed_similar + classact + Individualistic + Traditionalistic +
                Professionalization + unified + frailty(state_2),
                data = control_dataset)
```




### Frailty Models suing frailtypack

This is based off of Schorpp's comments.

```{r}
# mySurv <- Surv(time = dataset$case_year, time2 = dataset$Year, event = dataset$Override)
# 
# dataset2 <- dataset %>%
#   mutate(time = Year - case_year) %>%
#   dplyr::select(state_2, appointed, time, Override) %>%
#   na.omit()
# 
# frail1 <- frailtyPenal(Surv(time, Override) ~ cluster(state_2) + appointed, 
#                        data = dataset2, n.knots = 4, kappa = 50)
# 
# print(frail1)
# summary(frail1)
# plot(frail1)
```


```{r}
# dataset3 <- dataset %>%
#   mutate(time = Year - case_year) %>%
#   dplyr::select(state_2, appointed, time, Override, ideologically_similar) %>%
#   na.omit()
# 
# frail2 <- frailtyPenal(Surv(time, Override) ~ cluster(state_2) + appointed + ideologically_similar, 
#                        data = dataset3, n.knots = 4, kappa = 50)
```



```{r}
# dataset4 <- dataset %>%
#   mutate(time = Year - case_year) %>%
#   dplyr::select(state_2, appointed, time, Override, ideologically_similar, appointed_similar) %>%
#   na.omit()
# 
# frail3 <- frailtyPenal(Surv(time, Override) ~ cluster(state_2) + appointed + ideologically_similar +
#                          appointed_similar, 
#                        data = dataset4, n.knots = 4, kappa = 50)
```

```{r}
# summary(frail1)
# summary(frail2)
# summary(frail3)
```

```{r}
# print(frail1)
# print(frail2)
# print(frail3)
```

```{r}
# plot(frail1)
# plot(frail2)
# plot(frail3)
# 
# 
# plot.frail(frail2)
# frail1
```
