---
title: "Model"
author: "Richard G. Gardiner"
date: "2/20/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data and Packages

When you start looking at the model, look at [David's Tidy Tuesday project](https://mobile.twitter.com/drob/status/1082318029168566273?s=21). Also check out [this post](https://hockey-graphs.com/2019/01/07/quantifying-differences-between-the-regular-season-and-playoffs-using-survival-analysis/?blogsub=confirming#subscribe-blog).  Lastly, I like [this link](https://rpubs.com/daspringate/survival)

```{r, packages}
library(tidyverse)
library(survival)
library(survminer)
library(broom)
library(scales)
library(devtools)
library(ggfortify)
library(survivalAnalysis)
library(Amelia)
library(frailtypack)

theme_set(theme_light())
```

```{r}
dataset <- read_csv("data/all_years_all_variables.csv")
head(dataset)

dataset <- dataset %>%
  mutate(Year = Year + 0.01,
         Override = ifelse(Override == 109, 0, Override))



dataset %>%
  group_by(masterid) %>%
  count(masterid)
```

## Amelia 

This uses the data from Amelia to fill in missing data.  
```{r Amelia}
# filled_data <- amelia(dataset, idvars = c("state_abb", "State"))
# 
# write.amelia(filled_data, file.stem = "Amelia_data", format = "csv")
# 
# dataset <- read_csv("data/Amelia_data1.csv")
# 
# dataset
```


Here I am adding the absolute distance variables:

```{r}
# dataset$court_gov_dist <- abs(dataset$judicial - dataset$governor)
# dataset$court_lower_dist <- abs(dataset$judicial - dataset$lower)
# dataset$court_upper_dist <- abs(dataset$judicial - dataset$upper)
```



## Initial Analysis (not Amelia)

```{r}
# Model
election_cluster <- coxph(formula = Surv(case_year, Year, Override) ~ appointed + cluster(state_2),
                 data = dataset)
summary(election_cluster)

# graph prep
elections_cluster_fit <- survfit(formula = Surv(case_year, Year, Override) ~ appointed + cluster(state_2),
                 data = dataset)
# graph
tidy(elections_cluster_fit) %>%
  ggplot(aes(time, estimate, color = strata)) +
  geom_line() +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .2) +
  scale_color_discrete(name = "Selection System",
                       breaks = c("appointed=1", "appointed=0"),
                       labels = c("Appointed", "Elected")) +
  theme_light() +
  labs(y = "Estimated Likelihood of Survival",
       x = "Year")

# ggsave("figures/Selection Survival.jpeg")

cox.zph(election_cluster)
```

```{r}
ideology_cluster <- coxph(formula = Surv(case_year, Year, Override) ~ appointed +
                ideologically_similar + cluster(state_2),
                data = dataset)

summary(ideology_cluster)

ideology_cluster_fit <- survfit(formula = Surv(case_year, Year, Override) ~ appointed +
                ideologically_similar + cluster(state_2),
                data = dataset)


tidy(ideology_cluster_fit) %>%
  mutate(strata = if_else(strata == "appointed=0, ideologically_similar=1", "Elected and Similar",
                          if_else(strata == "appointed=0, ideologically_similar=0", "Elected and Dissimilar",
                                  if_else(strata == "appointed=1, ideologically_similar=0", 
                                          "Appointed and Dissimilar", "Appointed and Similar")))) %>%
  ggplot(aes(time, estimate, color = strata)) +
  geom_line() +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .2) +
  facet_wrap(~strata) +
  theme_light() +
  labs(y = "Estimated Likelihood of Survival",
       x = "Year",
       color = "Selection System and \nIdeological Similarity")

# ggsave("figures/Selection and Ideology survival.jpeg")

cox.zph(ideology_cluster)
```

```{r}
# creating interaction term
dataset <- dataset %>%
  mutate(appointed_similar = appointed * ideologically_similar)


interactions_cluster <- coxph(formula = Surv(case_year, Year, Override) ~ 
                                appointed + ideologically_similar + appointed_similar +
                                + cluster(state_2),
                data = dataset)
summary(interactions_cluster)

interactions_cluster_fit <- survfit(formula = Surv(case_year, Year, Override) ~ appointed +
                              ideologically_similar + appointed_similar + cluster(state_2),
                data = dataset)

tidy(interactions_cluster_fit) %>%
  mutate(strata = if_else(strata == "appointed=0, ideologically_similar=0, appointed_similar=0", 
                          "Elected and Dissimilar",
                          if_else(strata == "appointed=0, ideologically_similar=1, appointed_similar=0",
                                  "Elected and Similar",
                                  if_else(strata == 
                                            "appointed=1, ideologically_similar=1, appointed_similar=1", 
                                          "Appointed and Similar", "Appointed and Dissimilar")))) %>%
  filter(strata %in% c("Elected and Dissimilar", "Elected and Similar", "Appointed and Similar",
                       "Appointed and Dissimilar")) %>%
  ggplot(aes(time, estimate, color = strata)) +
  geom_line() +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .2) +
  facet_wrap(~strata) +
  theme_light() +
  labs(y = "Estimated Likelihood of Survival",
       x = "Year",
       color = "Interaction between Selection System \nand Ideological Similarity")  

# ggsave("figures/Interaction Survival.jpeg")

cox.zph(interactions_cluster)
```

```{r}
control_dataset <- dataset %>%
  filter(agency != 88) %>%
  mutate(Individualistic = ifelse(Traditionalistic == 0 & Moralistic == 0, 1, 0))

controls_cluster <- coxph(formula = Surv(case_year, Year, Override) ~ appointed +
                ideologically_similar + appointed_similar + classact + Individualistic + Traditionalistic +
                Professionalization + unified + cluster(state_2),
                data = control_dataset)

summary(controls_cluster)

controls_cluster_fit <- survfit(formula = Surv(case_year, Year, Override) ~ appointed +
                ideologically_similar + appointed_similar + classact + Individualistic + Traditionalistic +
                Professionalization + unified + cluster(state_2),
                data = control_dataset)

tidy(controls_cluster_fit) %>%
  mutate(strata = if_else(strata == "appointed=0, ideologically_similar=0, appointed_similar=0, classact=FALSE, Individualistic=0, Traditionalistic=0, Professionalization=1, unified=1", "elected, similar, not professional",
                           if_else(strata == "appointed=0, ideologically_similar=0, appointed_similar=0, classact=FALSE, Individualistic=0, Traditionalistic=0, Professionalization=5, unified=1", "elected, similar, most professional", strata))) %>%
  filter(strata %in% c("elected, similar, not professional", "elected, similar, most professional",
                       "appointed, similar, not professional", "appointed, similar, most professional")) %>%
  ggplot(aes(time, estimate, color = strata)) +
  geom_line() +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .2) +
  facet_wrap(~strata) +
  theme_light() +
  labs(y = "Estimated Likelihood of Survival",
       x = "Year",
       color = "Selection System and Professionalization")  

# ggsave("survival controls.jpeg")

cox.zph(controls_cluster)

print(tidy(controls_cluster_fit$strata))
```








## Using videos in attempt to better understand survival models

Creating survival object
```{r}
mySurv <- Surv(time = dataset$case_year, time2 = dataset$Year, event = dataset$Override)
head(mySurv) # plus sign means that it is centered
```


Baseline with no predictor
```{r}
fit <- survfit(mySurv ~ 1)

fit

plot(fit, ylim = c(0.9, 1), xlim = c(1995, 2013)) # kaplan meier curve
?plot
```

These Kaplan Meier tests/curves are nice, but have limitations.  It can only incorporate categorical variables.  It becomes difficult when you have many x variables (controlling becomes too hard), and there is no good way to summarize relationships (no betas) given that it is non-parametric
```{r}
states <- survfit(mySurv ~ dataset$State)

states

# kaplan meier curve
plot(fit, col = "red", lwd = 8, ylim = c(0.992, 1), xlim = c(1995, 2013), conf.int = FALSE) # baseline
lines(states) # state model


# trying to run a log rank test (like a chi-square test), but it is giving me an error
# survdiff(mySurv ~ dataset$state_abb) 
```

Elections
```{r}
elections <- survfit(mySurv ~ dataset$appointed)

plot(elections, ylim = c(0.998, 1), xlim = c(1995, 2013), col = c("blue", "red"))
```





### Cox proportional hazard ratio analysis

- Survival analysis allowing inclusing of multiple predictors
- Proportional Hazard assumption
    - assumes a constant hazard ratio over time
    - Groups compared diverge in the KM plot at the same rate (parallel)
- Coefficients represent the increase in the log Hazard Rate for one unit increase in predictor with all other covariates constant
- e^beta represents the Hazard rate for one unit increase in predictor with all others constant
- beta < 0 means that increasing x is associated with lower risk (hazard) and longer survival times
- to test hypothesis of beta = 0, we can use a Wald or LR test

```{r}
cox_states <- coxph(mySurv ~ dataset$State)

summary(cox_states)

tidy(cox_states) %>%
  mutate(term = str_remove_all(term, "dataset\\$State"),
         estimate2 = exp(estimate),
         significance = ifelse(estimate > 0 & p.value < 0.05, "positive",
                               ifelse(estimate > 0 & p.value < 0.05, "negative",
                                      "not significant")),
         term = fct_reorder(term, estimate)) %>%
  # filter(p.value < 0.05) %>%
  filter(term != "New Mexico") %>%
  ggplot(aes(x = term, y = estimate, color = significance)) +
  geom_point() +
  geom_pointrange(aes(ymin = conf.low, ymax = conf.high)) +
  scale_color_manual(values = c("gray60", "green2")) +
  coord_flip() +
  theme_minimal()

tidy(cox_states) %>%
  mutate(term = str_remove_all(term, "dataset\\$State"),
         estimate2 = exp(estimate),
         significance = ifelse(estimate > 0 & p.value < 0.05, "positive",
                               ifelse(estimate > 0 & p.value < 0.05, "negative",
                                      "not significant")),
         term = fct_reorder(term, conf.low)) %>%
  filter(p.value < 0.05) %>%
  filter(term != "New Mexico") %>%
  ggplot(aes(x = term, y = estimate)) +
  geom_point() +
  geom_pointrange(aes(ymin = conf.low, ymax = conf.high)) +
  scale_color_manual(values = c("green2")) +
  coord_flip() +
  theme_minimal()
```


Hazard is the probability that you die now, given that you are alive (probability of being overriden, given that you haven't been overridden right now).

A hazard ratio is just like an odds-ratio.  The cox proportional can't predict the hazard for a given person with a given set of values at a given moment.  We can only COMPARE people, because we don't know the value of the baseline hazard


WHAT IF I SHOULD BE SUBTRACTING THE YEAR? ehh.. Really doesn't matter
```{r}
mySurv2 <- Surv(time = (dataset$case_year - 1994), time2 = (dataset$Year - 1994), event = dataset$Override)

fitTest <- survfit(mySurv2 ~ 1)

fitTest

plot(fitTest) # kaplan meier curve
```











## Frailty using the survival package (MAIN RESULTS NOW)

```{r}
frailty1 <- coxph(Surv(time = (dataset$case_year - 1994), time2 = (dataset$Year - 1994), event = dataset$Override) ~ dataset$appointed + frailty(dataset$state_abb))


surv_frailty1 <- survfit(Surv(time = (dataset$case_year - 1994), 
                              time2 = (dataset$Year - 1994), 
                              event = dataset$Override) ~ 
                           dataset$appointed + frailty(dataset$state_abb))



surv_frailty1 %>%
  tidy() %>%
  ggplot(aes(time, estimate, color = strata)) +
  geom_line(show.legend = FALSE) +
  labs(x = "Time", y = "Estimate")
```

```{r}
frailty2 <- coxph(Surv(time = (dataset$case_year - 1994), 
                              time2 = (dataset$Year - 1994), 
                              event = dataset$Override) ~ 
                           dataset$appointed + dataset$ideologically_similar +
                           frailty(dataset$state_abb))

summary(frailty2, type = "text")

surv_frailty2 <- survfit(Surv(time = (dataset$case_year - 1994), 
                              time2 = (dataset$Year - 1994), 
                              event = dataset$Override) ~ 
                           dataset$appointed + dataset$ideologically_similar +
                           frailty(dataset$state_abb))

frailty2_data <- surv_frailty2 %>%
  tidy() %>% 
  mutate(appointed = ifelse(str_detect(strata, "appointed=0") == TRUE,
                            "Elected", "Appointed")) %>%
  group_by(time, appointed) %>%
  summarise(estimate = mean(estimate),
            conf.high = mean(conf.high),
            conf.low = mean(conf.low))

frailty2_data %>%
  ggplot(aes(time, estimate, color = appointed)) +
  geom_line() +
  labs(color = "Appointed", x = "Time", y = "Estimate")

surv_frailty2 %>%
  tidy() %>% 
  mutate(appointed = ifelse(str_detect(strata, "appointed=0") == TRUE,
                            "Elected", "Appointed")) %>%
  ggplot(aes(time, estimate, color = appointed)) +
  geom_line(aes(group = strata)) +
  geom_line(data = frailty2_data, lwd = 1, color = "gray60") +
  labs(x = "Time", y = "Estimate", color = "Selection System",
       caption = "Gray line is the group average") +
  facet_wrap(~appointed) 
```


```{r}
frailty3 <- coxph(Surv(time = (dataset$case_year - 1994), 
                              time2 = (dataset$Year - 1994), 
                              event = dataset$Override) ~ 
                           dataset$appointed + dataset$ideologically_similar +
                           dataset$appointed_similar +
                           frailty(dataset$state_abb))

summary(frailty3)

surv_frailty3 <- survfit(Surv(time = (dataset$case_year - 1994), 
                              time2 = (dataset$Year - 1994), 
                              event = dataset$Override) ~ 
                           dataset$appointed + dataset$ideologically_similar +
                           dataset$appointed_similar +
                           frailty(dataset$state_abb))

surv3_data <- surv_frailty3 %>%
  tidy() %>% 
  mutate(appointed = ifelse(str_detect(strata, "appointed=0") == TRUE,
                            "Elected", "Appointed"),
         ideological_dissimilar = ifelse(str_detect(strata, "ideologically_similar=0") == TRUE,
                                         "Dissimilar", "Similar")) %>%
  group_by(time, appointed, ideological_dissimilar) %>%
  summarise(estimate = mean(estimate),
            conf.high = mean(conf.high),
            conf.low = mean(conf.low)) %>%
  mutate(final = paste(appointed, ideological_dissimilar))
  

surv3_data %>%
  ggplot(aes(time, estimate, color = appointed)) +
  geom_line() +
  labs(color = "Appointed", x = "Time", y = "Estimate") +
  facet_wrap(~ideological_dissimilar)



surv_frailty3 %>%
  tidy() %>% 
  mutate(appointed = ifelse(str_detect(strata, "appointed=0") == TRUE,
                            "Elected", "Appointed"),
         ideological_dissimilar = ifelse(str_detect(strata, "ideologically_similar=0") == TRUE,
                                         "Dissimilar", "Similar"),
         final = paste(appointed, ideological_dissimilar)) %>%
  ggplot(aes(time, estimate, color = final)) +
  geom_line(aes(group = strata)) +
  geom_line(data = surv3_data, lwd = 1, color = "gray60") +
  labs(x = "Time", y = "Estimate", color = "Selection System",
       caption = "Gray line is the group average") +
  facet_wrap(~final) 
```



```{r}
control_dataset <- dataset %>%
  filter(agency != 88) %>%
  mutate(Individualistic = ifelse(Traditionalistic == 0 & Moralistic == 0, 1, 0))

controls_cluster_frail <- coxph(formula = Surv(case_year, Year, Override) ~ appointed +
                ideologically_similar + appointed_similar + classact + Individualistic + Traditionalistic +
                Professionalization + unified + frailty(state_2),
                data = control_dataset)

summary(controls_cluster_frail)

controls_cluster_fit_frail <- survfit(formula = Surv(case_year, Year, Override) ~ appointed +
                ideologically_similar + appointed_similar + classact + Individualistic + Traditionalistic +
                Professionalization + unified + frailty(state_2),
                data = control_dataset)
```




### Frailty Models suing frailtypack

This is based off of Schorpp's comments.

```{r}
# mySurv <- Surv(time = dataset$case_year, time2 = dataset$Year, event = dataset$Override)
# 
# dataset2 <- dataset %>%
#   mutate(time = Year - case_year) %>%
#   dplyr::select(state_2, appointed, time, Override) %>%
#   na.omit()
# 
# frail1 <- frailtyPenal(Surv(time, Override) ~ cluster(state_2) + appointed, 
#                        data = dataset2, n.knots = 4, kappa = 50)
# 
# print(frail1)
# summary(frail1)
# plot(frail1)
```


```{r}
# dataset3 <- dataset %>%
#   mutate(time = Year - case_year) %>%
#   dplyr::select(state_2, appointed, time, Override, ideologically_similar) %>%
#   na.omit()
# 
# frail2 <- frailtyPenal(Surv(time, Override) ~ cluster(state_2) + appointed + ideologically_similar, 
#                        data = dataset3, n.knots = 4, kappa = 50)
```



```{r}
# dataset4 <- dataset %>%
#   mutate(time = Year - case_year) %>%
#   dplyr::select(state_2, appointed, time, Override, ideologically_similar, appointed_similar) %>%
#   na.omit()
# 
# frail3 <- frailtyPenal(Surv(time, Override) ~ cluster(state_2) + appointed + ideologically_similar +
#                          appointed_similar, 
#                        data = dataset4, n.knots = 4, kappa = 50)
```

```{r}
# summary(frail1)
# summary(frail2)
# summary(frail3)
```

```{r}
# print(frail1)
# print(frail2)
# print(frail3)
```

```{r}
# plot(frail1)
# plot(frail2)
# plot(frail3)
# 
# 
# plot.frail(frail2)
# frail1
```
